A brief comparison of speed of Interior Point and EM algorithms

### 2016-10-19 analysis/IPvsEM.Rmd
IPvsEM.Rmd runs correctly; made a few minor adjustments to code and text.

```{r model_fitting}
df = data.frame()
cat("method --optmethod-- sim\n")
for(method in c("fdr","shrink")) {
  for(optmethod in c("mixIP","cxxMixSquarem")) {
    df = rbind(df,data.frame(method = method,optmethod = optmethod,
           timed_sims(list(method = method,optmethod = optmethod),
                           nsim = 50,nmin = 10,nmax = 100000)))
  }
}
cat("\n")
```

@@

```{r plot_results1}
qplot(x = n,y = elapsed.time,data = df,col=optmethod,facets=.~method,ylab="elapsed time (s)")
```

And a zoom-in on "small" problems with $n<5000$.

```{r plot_results2}
ggplot2::qplot(x=n,y=elapsed.time,data=dplyr::filter(df,n<5000),col=optmethod,facets=.~method,ylab="elapsed time (s)")
```

@@

However, for small problems (n<5000) the EM is adequate for many practical purposes, solving within a few seconds. This is particularly true for the penalty term (method=fdr), which
 helps the EM converge, presumably because it is helping identifiability, removing the large flat parts of the likelihood objective function.

Indeed, for small problems with the penalty term (n<2000) EM with squarem acceleration is a little faster in these comparisons than IP.

>>>>

```{r model_fitting}
df = data.frame()
cat("method --optmethod-- sim\n")
for(method in c("fdr","shrink")) {
  for(optmethod in c("mixIP","cxxMixSquarem")) {
    df = rbind(df,data.frame(method = method,optmethod = optmethod,
           timed_sims(list(method = method,optmethod = optmethod),
                           nsim = 50,nmin = 10,nmax = 100000)))
  }
}
cat("\n")

# Summarize the results in a table.
cat("Average computation time (seconds):\n")
print(as.table(by(df,df[c("method","optmethod")],
                  function (x) mean(x$elapsed.time))))
```

@@

```{r plot_results}
qplot(x = n,y = elapsed.time,data = df,col = optmethod,facets = .~method,
      ylab = "elapsed time (s)")
```

Zoom-in on "small" problems with $n < 5000$.

```{r plot_results_magnified}
qplot(x = n,y = elapsed.time, data = subset(df,n < 5000),
      col = optmethod,facets = .~method,ylab = "elapsed time (s)")
```

@@

However, for small problems (n < 5000) the EM is adequate for many practical
purposes, solving within a few seconds. This is particularly true for the
penalty term (method = fdr), which helps the EM converge, presumably because
it is helping identifiability, removing the large flat parts of the likelihood
objective function.

Indeed, for small problems with the penalty term (n < 2000) EM with squarem
acceleration is a little faster in these comparisons than IP.
